Stack Safety as a Security Property

### 2020-07-08 main.tex
Filling in policy discussion and more tweaks where needed

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\input{macros}
\input{examples}

@@

\subsection{Unsafe Variations}
\sna{Variations here collides with state variations. I propose substitutions.}
\leo{I agree there is a collision, not sure about ``substitutions''}

The execution trace shown in \cref{fig:running-trace-b} is an example
of a ``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. In the rest of
this section we'll go over examples of unsafe behaviors, before
formally capturing what stack-safe means formally and discussing
enforcement techniques from the literature in later sections.

@@

\bcp{More explanation needed, plus examples.}A policy is a general model of an enforcement mechanism, consisting of
a set of policy states \(\pol \in \POLS\), and a policy step function
\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\).

@@

\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
            {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}

\leo{Notation: Should $\stepstoobs{}$ use harpoon for the lifted
  step?}\bcp{Yes.}

@@

\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states $\POLS$ is the unit type, and its
policy step function is the constant function that produces the unit, $\lambda
\_ . \top$, thereby allowing all calls the machine step function. This
corresponds to the lifting of our machine model to the policy setting.

\paragraph*{Software-only policies}
%
\rb{TODO In a way, in these schemes the policy state is built in as part of the
  program state and transparent to the hardware, although it could be simulated
  by an overlay where enforcement is represented at the policy level.}

@@

\paragraph*{Micropolicies}
%
\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
  policy? }

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a code map $\codemap$, an $n$-{\em transition} a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\input{macros}
\input{examples}

@@

\subsection{Unsafe Variations}
\sna{Variations here collides with state variations. I propose substitutions.}
\leo{I agree there is a collision, not sure about ``substitutions''}
\sna{Mutations, drawing from the randomized testing literature?}

The execution trace shown in \cref{fig:running-trace-b} is an example
of a ``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. In the rest of
this section we'll go over examples of unsafe behaviors, before
formally capturing what stack-safe means formally and discussing
enforcement techniques from the literature in later sections. Throughout
the paper examples will be localize modifications of the example, by
substituting an instruction at a given PC, written:
\[\PCname :\mathit{inst}\]

@@

\bcp{More explanation needed, plus examples.}
A policy is a general model of an enforcement mechanism, consisting of a
set of policy states \(\pol \in \POLS\), a policy step function
\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
\POLS\). The initializer establishes the initial policy state for a given
machine state, but is partial to reflect that some policies are only defined
on states that are well-formed in some way.

@@

\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
               \polstep \pol_2\)}
            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
               \mpstate{\mach_2}{\pol_2}\)}

@@

\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states $\POLS$ is the unit type, and its
policy step function is the constant function that produces the unit, $\lambda
\_ . \top$, thereby allowing all calls the machine step function. Its
initialization function is defined on all states. This corresponds to the
lifting of our machine model to the policy setting. Machines that fulfill a
property by construction will also fulfill that property with the null
policy..

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler
to (e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
canaries \citep{Cowan+98}. These approaches are represented as policies that
accept only initial machine states that are well-formed for the policy (due
to being produced by the compiler.) The set of policy states \(\POLS\) is again
the unit type and the step function constant. Property enforcement comes from
the fact that only ill-formed states can induce a trace that violates the
property.

\rb{TODO In a way, in these schemes the policy state is built in as part of the
  program state and transparent to the hardware, although it could be simulated
  by an overlay where enforcement is represented at the policy level.}

@@

\paragraph*{Micropolicies}
%
Micropolicies are another major family of policies \citep{pump_oakland2015},
which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
Here the machine state is extended with tags paired with all components of the
system, and a monitor checks for each operation that the tags on the operands
obey a set of rules. One could treat tags directly as part of the machine
state, and build the rules into the step function. But local state
encapsulation properties as we frame them are not intuitive in their
interaction with tags or other additional enforcement state, so it is helpful
to separate tags for clarity. This also distinguishes a hardware fail-stop from
a software fail state.
\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
  policy? }

\paragraph*{Machine Termination}

In our model, the machine step function is total -- a program does not halt
on its own, but may instead enter a silent terminal loop. Runs are only cut
short by the policy step function. This means that software-only policies that
can catch errors must be modeled as reaching an error state, not outright
halting.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a code map $\codemap$, an $n$-{\em transition} a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}
